M\"obius Inversion: Construction and Applications

### 2017-04-10 Drafts/Draft.tex
Figures

\documentclass[12pt]{pom_thesis}
\usepackage[T1]{fontenc}
\input{commands_packages}
\usetikzlibrary{decorations.pathreplacing,angles,quotes}
\usepackage{mathrsfs}
\usepackage[euler-digits]{eulervm} %TODO: Maybe change this
\author{Leo Selker}
\advisor{Vin de Silva}
\title{M\"obius Inversion: Construction and Applications}
\newcommand{\fix}{\text{fix }}
\newcommand{\cl}{\text{cl }}
\DeclareMathOperator{\cat}{cat}
\DeclareMathOperator{\ps}{\mathscr{P}}
\DeclareMathOperator{\sub}{Sub}
\DeclareMathOperator{\stab}{Stab}
\DeclareMathOperator{\orb}{Orb}
\DeclareMathOperator{\fixx}{fix}
\DeclareMathOperator{\obj}{ob}
\DeclareMathOperator{\im}{Im}
\newcommand{\catname}[1]{\textbf{\textsc{#1}}}

@@

\newpage
\pagenumbering{arabic}
\begin{chapter}{Introduction to M\"obius Inversion}\label{chap_intro}
\section{Introduction and Motivation}
%TODO: Expand introduction with more explicit examples
Suppose that we have the integer sequence $g$ of ones:
\[
g : 1, 1, 1, \cdots
\]
We can let $f$ be the sequence of cumulative sums of $g$. So $f_k = g_1 + \cdots + g_k$. We can write down the values:
\[
f: 1,2,3,\cdots
\]
In general, summing up values is easy. Finding values of $f$ from $g$ was not difficult.

@@

\section{Definition of M\"obius Inversion}
We are now ready to define the M\"obius function on partially ordered sets.
\begin{defn}\label{def_mob_pos}
Let $X$ be a lower finite poset. Let $\zeta$ be the indicator function defined above. Let $\mu = \zeta^{-1}$. Then $\mu$ is the \emph{M\"obius function} on $X$.
\end{defn}
\begin{rmk}
Recall that $\zeta^{-1}$ is the matrix inverse rather than the functional inverse, and that it exists by \ref{inv_ex}.
\end{rmk}
\begin{rmk}\label{sum_form}
We can unpack the linear algebra to get a definition analagous to \eqref{sum_form_intro}: We can say that $\mu$ is the unique function from $X\times X$ to $\Rr$ such that, for all $x,y \in X$,
\[
\sum_{z \in X} \mu(x,z)\zeta(z,y) = \delta(x,y) = \sum_{z \in X} \zeta(x,z)\mu(z,y),
\]
or
\[
\sum_{z \leq y} \mu(x,z) = \delta(x,y) = \sum_{z \geq x} \mu(z,y).
\]
\end{rmk}

To complete the definition of M\"obius inversion, we turn to the standard use of this function. 
\begin{thm}[Single Variable M\"obius Inversion]\label{mob_inv}
Let $X$ be a lower finite poset. Let $f,g:X \rightarrow \Rr$ be given. Let $\zeta:X \times X \rightarrow \Rr$ be defined as usual, and let $\mu = \zeta^{-1}$. Then:
\begin{align*}
f = g \zeta & \iff g = f \mu.
\end{align*}
Equivalently,
\begin{align*}
\forall x \in X, f(x) = \sum_{y \in X} g(y)\zeta(y,x) &\iff \forall x \in X, g(x) = \sum_{y \in X} f(y)\mu(y,x).
\end{align*}
\end{thm}
\begin{proof}
Recall that by \ref{inv_ex}, $\mu = \zeta^{-1}$ exists. Given that, the result follows if we right-multiply both sides by $\mu$. The first statement becomes the second when the sums are made explicit.
\end{proof}

This result is quite straightforward based on how we defined $\zeta$ and $\mu$. In fact, it feels strange that a supposedly central fact is just the inversion of a single matrix operation. In fact, the simplicity of the inversion step is the whole point of M\"obius inversion. The tricky part is formulating a given problem such that \ref{mob_inv} can be applied to it. We will see many cases where values of $f$ from the theorem are readily attainable, but values of $g$ are what we really want. The M\"obius function allows us to obtain those values. This procedure - the use of the M\"obius function to reverse the process of summing terms as in theorem \ref{mob_inv} - is what is referred to as M\"obius inversion.

@@

In this case, if $\leq$ represents the poset relation from \ref{lattice}, then $m \leq n$ if, for each prime $p$, $m$'s factorization contains no more copies of $p$ than $n$'s factorization. This is the same as saying that $m|n$. So the poset relation becomes divisibility. We will explore this example more in the next sections.
\end{examp}
It turns out that several important posets have this lattice structure. So the goal is to determine the M\"obius function on certain subsets of the integer lattice, and then transfer those results to other partially ordered sets. To do so, we first need a lemma which applies to the M\"obius function on any poset.
\begin{lemma}\label{recurse}
Let $X$ be a lower finite partially ordered set, and let $\zeta, \mu$ be defined as usual. Then for all $a, b \in X$,
\[
\mu(a,b) = 
\begin{cases}
0 & a \nleq b\\
\delta(a,b)-\sum_{c \in [a,b)}\mu(a,c) & else.
\end{cases}
\]
\end{lemma}
\begin{rmk}
Note that the piecewise form is actually not necessary, because if $a \nleq b$ then the sum will be empty. So it is correct - albeit less clear- to just write
\[
\mu(a,b) = \delta(a,b)-\sum_{c \in [a,b)}\mu(a,c).
\]
\end{rmk}
\begin{proof}
We fix $a,b$, and then rearrange terms starting from \ref{sum_form}:
\begin{align*}
\sum_{c \in X}\mu(a,c) \zeta(c,b) &= \delta(a,b)\\
\sum_{c \in [a,b)} \mu(a,c) + \mu(a,b) &= \delta(a,b)\\
\mu(a,b) &= \delta(a,b) - \sum_{c \in [a,b)} \mu(a,c) 
\end{align*}
\end{proof}
We are now ready to write down $\mu$ on certain subsets of the integer lattice. In order to ensure that the sums involved are predictable, we need an additional condition on the subsets we consider:

\begin{lemma}\label{int_lattice}
Let $L$ be the $n$-dimensional integer lattice, with the usual poset relation. Let $\zeta, \mu$ be defined as usual on $L$. Let $x=(x_1,\dots,x_n),y = (y_1,\dots,y_n)$ be elements of $L$. Then:
\[
\mu(x,y) =
\begin{cases}
0 & x\nleq y\\
0 & \exists i \in \Nn: y_i-x_i \geq 2\\
(-1)^{\sum_{i=1}^ny_i-x_i} & \text{otherwise}
\end{cases}
\]
\end{lemma}

\begin{proof}
If $x \nleq y$, then by \ref{recurse}, $\mu(x,y) = 0$. This is the first case.

We now proceed by induction on $\sum_{i=1}^ny_i-x_i$. If $x \leq y$ and $\sum_{i=1}^ny_i-x_i = 0$, then $x = y$. So by \ref{recurse}, $\mu(x,y) = 1$. This is the base case.

Suppose that we have \ref{int_lattice} for any $x \leq y$ where $\sum_{i=1}^ny_i-x_i \leq k-1$. Let $x \leq y$ be fixed such that $\sum_{i=1}^ny_i-x_i =k$. Suppose that for some $j \in \Nn, y_j-x_j \geq 2$. Let $y' = (y_1,\dots,y_j-1,\dots,y_n)$. Note that $x \leq y' < y$. So we can decompose the sum from $\ref{int_lattice}$. Let
\begin{align*}
\alpha &= \sum_{x \leq z \leq y'} \mu(x,z)\\
\beta &= \sum_{y' < z < y}\mu(x,z).\\
\end{align*}
Then
\begin{align*}
\mu(x,y) &= -\sum_{x \leq z \leq y}\mu(x,z)\\
&= -\sum_{x \leq z \leq y'} \mu(x,z) - \sum_{y' < z < y}\mu(x,z)\\
&= -(\alpha + \beta).
\end{align*}
We note that
\[\alpha = \sum_{z \in L} \mu(x,z)\zeta(z,y').
\]
So by \ref{sum_form}, $\alpha=0$.

@@

Finally, suppose that for all $i = 1,\dots,n, y_i-x_i \leq 1$. Because $\sum_{i=1}^ny_i-x_i =k$, there must be $k$ indices where $x$ and $y$ differ by one, while at the other $n-k$ indices they are equal. So for each $j = 0,\dots,k-1$, there are $k \choose j$ elements $z$ of $X$ where $x \leq z < y$ and $\sum_{i = 1}^n y_i-z_i = j$. For each of these, by the inductive hypothesis, $\mu(x,z) = (-1)^j$. So, using \ref{sum_form}, we can write
\begin{align*}
\mu(x,y)& = -\sum_{j = 1}^k {k \choose j} (-1)^j\\
&= (-1)^k.
\end{align*}
\end{proof}
This result will be powerful for determining values of $\mu$ on various posets whose structure is similar to an integer lattice. 
\begin{thm}\label{mapping}
Let $(X, \leq_X),(Y, \leq_Y)$ be lower finite posets. Let $x_1 \leq x_2 \in S$ be given. Suppose that there exists a bijective map
\[
\phi: [x_1, x_2] \leftrightarrow [\phi(x_1), \phi(x_2)]
\]
such that for all $z_1, z_2 \in [x_1, x_2]$,
\[ 
z_1 \leq_X z_2 \iff \phi(z_1) \leq_Y \phi(z_2).
\] 
Let $\mu_X, \mu_Y$ denote the M\"obius functions on $X$ and $Y$ respectively. Then:
\[
\mu_X(x_1, x_2) = \mu_Y(\phi(x_1), \phi(x_2)).
\]
\end{thm}
This theorem tells us that if we can find a portion of a familiar poset which has identical structure to a portion of an unknown poset, we can transfer what we know about values of $\mu$ on that portion from one to the other.
\begin{proof}
We induct on $\#[x_1, x_2]$. Because $x_1 \leq x_2$, if $[x_1, x_2] = 1$, then $x_1 = x_2$. So by \ref{recurse}, $\mu_X(x_1, x_2) = 1$. Because $\phi(x_1) = \phi(x_2), \mu_Y(\phi(x_1), \phi(x)2) = 1$ also.

Suppose the theorem is true if $\#[x_1, x_2] \leq k-1$ for some integer $k$. Fix $x_1, x_2$ such that $\#[x_1, x_2] = k$. Then by \ref{recurse},
\[
\mu_Y(\phi(x_1), \phi(x_2)) = -\sum_{y \in [\phi(x_1), \phi(x_2)]}\mu_Y(\phi(x_1), y)
\]
Because $\phi$ is an isomorphism which preserves $\leq$, we can rewrite this sum:
\[
\mu_Y(\phi(x_1), \phi(x_2)) = -\sum_{x \in [x_1, x_2]}\mu_Y(\phi(x_1), \phi(x))
\]
We can restrict $\phi$ to $[\phi(x_1), \phi(x)]$ to satisfy the hypotheses for all intermediate elements $x$. So we can apply the inductive hypothesis to all the terms of this sum:
\begin{align*}
\mu_Y(\phi(x_1), \phi(x_2)) &= -\sum_{x \in [x_1, x_2]}\mu_X(x_1, x)\\
&= \mu_X(x_1, x_2)
\end{align*}
\end{proof}
The following corollary follows immediately if we let $\phi$ be the inclusion map:
\begin{cor}\label{subset}
Let $X$ be a poset. Let $I$ be an interval in $X$. In other words, let $I = [z_1, z_2]$ for some $z_1, z_2 \in X$. Then
\[
\mu_I = \mu_X|_I.
\]
\end{cor}
In the next section, we explore a poset where this fact will be extremely useful.

@@

We will consider the poset $(\Zz^{+}, |)$. This is the set of positive integers ordered by divisibility (recall that $d|n$ if $d$ divides $n$ without remainder). Note that this poset is lower finite: For any number $n$, the set $\{q \in \Zz^{+}: q|n\} \subseteq \{q \in \Zz^{+}: q \leq n\}$, and the latter set contains $n$ elements, which is finite (note that $d|n$ means that $d$ divides $n$ without remainder). This means we are able to use M\"obius inversion. Rather than compute the M\"obius function directly, we can use \ref{int_lattice} and \ref{mapping} to reuse what we already know.

@@

\begin{proof}
Let $d,n \in \Zz^+$ be given. We set up the isomorphism required for \ref{mapping}: Index the primes $p_1, p_2, \dots$. So $p_1=2, p_2=3$, and so on. Let $k$ be the index of the largest prime which divides either $d$ or $n$. Then let $\phi : [d,n] \rightarrow \Nn^k$ be defined by
\[
 p_1^{x_1}p_2^{x_2}\cdots p_k^{x_k} \mapsto (x_1, \cdots, x_k).
\]
We claim that $\phi$ is a bijection from $[d,n]$ to $\im(\phi)$.

@@

This allows us to use \ref{mapping}, and the result immediately follows.
\end{proof}

This definition of $\mu$ depends only on $\frac nd$, which means that we can abuse notation to write $\mu(\frac nd)$ instead of $\mu(d,n)$. This allows us to use \ref{mob_inv} to arrive at \eqref{init_mob}:
\[
f(n) = \sum_{d | n} g(d) \iff g(n) = \sum_{d | n}  f(d)\mu\left(\frac nd\right).
\]
\section{Euler's Totient} %TODO: Fix \Zz^+
Euler's Totient is a famous function in number theory, which counts the how many numbers are relatively prime to a particular value. We denote the function by $\varphi : \Nn \rightarrow \Nn$, and its definition is:
\[
\varphi(n) = \#\{1 \leq m \leq n : (m,n) = 1\},
\]
where $(m,n)$ denotes the greatest common divisor of $m,n$.

@@

We start with an important fact:
\begin{thm}[Gauss]\label{gauss}
Let $n \in \Nn$ be given. Then 
\[
n = \sum_{d|n}\varphi(d).
\]
\end{thm}
There are a variety of ways of proving this fact. Here we include a concise proof using group theory.
\begin{proof}
For any integer $d$, $\varphi(d)$ counts the generators of $C_d$. Each element of $C_n$ generates exactly one cyclic subgroup, so counting the total number of generators of cyclic subgroups of $C_n$ is equivalent to counting the elements of $C_n$. Also, $C_n$ has exactly one cyclic subgroup for each divisor of $n$ (and no others). So the sum indeed counts the number of elements of $C_n$, of which there are $n$.
\end{proof}
This sum gives us a value we have access to ($n$) in terms of a sum of values we want (values of $\varphi$). This is a perfect situation to apply M\"obius inversion, and undo that sum. We apply \eqref{init_mob}, which tells us
\[
\varphi(n) = \sum_{d|n} d\mu\left(\frac nd\right).
\]
Recall that \ref{mu_div} tells us how to write down values of $\mu$. Depending on what information we have access to, this might be much easier than just counting the number of values relatively prime to $n$. This is a straightforward example of how M\"obius inversion's ability to reverse a sum can be useful.
\end{chapter}

\begin{chapter}{Groups and Orbits} 
\section{Background}
Suppose that a group $G$ acts on a set $\Omega$.

I will use $G(x)$ to denote $x$'s orbit under $G$, and $\orb(\Omega, G)$ to denote the set of orbits in $\Omega$ under $G$'s action on it. 
\section{M\"obius Inversion on Poset of Subgroups}
We want to apply M\"obius inversion to finite groups. This means we will need a poset. We will use the \emph{poset of subgroups} (or \emph{lattice of subgroups}):
\begin{defn}
Let $G$ be a finite group. Then let the \textit{poset of subgroups} of $G$, or $\sub(G)$, be the poset consisting of subgroups of $G$ \emph{reverse} ordered by inclusion. In other words, we say that $A \leq B$ if $A \supseteq B$.
\end{defn}
\begin{rmk}
Because $G$ is finite, $\sub(G)$ is finite (so lower finite).
\end{rmk}
\begin{rmk}
We are ordering the poset this way for reasons that will soon become clear. But to reduce confusion, we will continue to use inclusion notation to define $\zeta$, and so $\mu$, rather than using $\leq$. Also note that I will be using $\subseteq$ rather than $\leq$ to denote subgroups, also to reduce confusion with the poset relation.
\end{rmk}
This puts us in the position to apply M\"obius inversion. We first define the usual functions on $\sub(G) \times \sub(G)$:
\begin{itemize}
\item Given a group $G$, let $\zeta:\sub(G) \times \sub(G) \rightarrow \Rr$ be defined by
\[
\zeta(a,b) = \begin{cases}
1 & A \supseteq B\\
0 & else
\end{cases}
\]
\item Let $\mu = \zeta^{-1}$, which exists by \ref{inv_ex}.
\end{itemize}
Recall that we manipulate $\zeta, \mu$ as if they were matrices.

Now, \ref{mob_inv} tells us that, for any row vectors $f,g$ of length $|\sub{G}|$, 
\[
f = g\zeta \iff g = f\mu.
\]
Now, of course, the challenge is to determine how to formulate the problem such that inverting $\zeta$ will be useful.
\section{Counting Orbits}
Throughout this section, let $G$ be a finite group acting on a set $\Omega$.

We start by defining some functions/vectors, each with entries corresponding with subgroups of $G$, which we will be using. Let:
\begin{align*}
f(A) &= \#\{x \in \Sigma: \stab(x) \supseteq A\} = |\fix(A)|\\
g(A) &= \#\{x \in \Sigma: \stab(x) = A\}
\sigma(A) &= |A|
\end{align*}

What now? The goal is to count the orbits in $\Omega$ under the action by $G$ - in other words, to find $|\orb(\Omega, G)|$. We will start with the assumption that we are able to find $\sigma$, which is true if we know the structure of $\sub(G)$. We will also assume that we are able to carry out the group action, thereby determining which elements of $\Omega$ are fixed by given subgroups of $G$. This gives us access to $f$. Then there are two steps we will follow:
\begin{description}
\item[Step 1:] Determine how to use $g$ and $\sigma$ to find $|\orb(\Omega, G)|$; and
\item[Step 2:] Find $g$ from $f$ using M\"obius inversion.
\end{description}

>>>>

\documentclass[12pt]{pom_thesis}
\usepackage[T1]{fontenc}
\input{commands_packages}
\usetikzlibrary{decorations.pathreplacing,angles,quotes}
\usepackage{mathrsfs}
\usepackage[euler-digits]{eulervm} 
\usepackage[font={small}]{caption}
\usepackage{float}
\author{Leo Selker}
\advisor{Vin de Silva}
\title{M\"obius Inversion: Construction and Applications}
\newcommand{\kron}{\boldsymbol{\otimes}}
\DeclareMathOperator{\fix}{Fix}
\DeclareMathOperator{\cat}{cat}
\DeclareMathOperator{\ps}{\mathscr{P}}
\DeclareMathOperator{\sub}{Sub}
\DeclareMathOperator{\stab}{Stab}
\DeclareMathOperator{\orb}{\mathcal{O}}
\DeclareMathOperator{\fixx}{fix}
\DeclareMathOperator{\obj}{ob}
\DeclareMathOperator{\im}{Im}
\newcommand{\catname}[1]{\textbf{\textsc{#1}}}

@@

\newpage
\pagenumbering{arabic}
\begin{chapter}{Introduction to M\"obius Inversion}\label{chap_intro}
\section{Introduction and Motivation}
%TODO: Capitalize "theorem, lemma,..."
%TODO: Expand introduction with more explicit examples
Suppose that we have the integer sequence $g$ of ones:
\[
g : 1, 1, 1, \cdots
\]
We can let $f$ be the sequence of cumulative sums of $g$. So $f_k = g_1 + \cdots + g_k$. We can write down the values:
\[
f: 1,2,3,\cdots
\]
In general, summing up values is easy. Finding values of $f$ from $g$ was not difficult.

@@

\section{Definition of M\"obius Inversion}
We are now ready to define the M\"obius function on partially ordered sets.
\begin{defn}\label{def_mob_pos}
Let $X$ be a lower finite poset. Let $\zeta$ be the indicator function defined above. Let $\mu = \zeta^{-1}$. Then $\mu$ is the \emph{M\"obius function} on $X$.
\end{defn}
\begin{rmk}
Recall that $\zeta^{-1}$ is the matrix inverse rather than the functional inverse, and that it exists by Lemma \ref{inv_ex}.
\end{rmk}
\begin{rmk}\label{sum_form}
We can unpack the linear algebra to get a definition analagous to \eqref{sum_form_intro}: We can say that $\mu$ is the unique function from $X\times X$ to $\Rr$ such that, for all $x,y \in X$,
\[
\sum_{z \in X} \mu(x,z)\zeta(z,y) = \delta(x,y) = \sum_{z \in X} \zeta(x,z)\mu(z,y),
\]
or
\[
\sum_{z \leq y} \mu(x,z) = \delta(x,y) = \sum_{z \geq x} \mu(z,y).
\]
\end{rmk}

To complete the definition of M\"obius inversion, we turn to the standard use of this function. 
\begin{thm}[Single Variable M\"obius Inversion]\label{mob_inv}
Let $X$ be a lower finite poset. Let $f,g:X \rightarrow \Rr$ be given. Let $\zeta:X \times X \rightarrow \Rr$ be defined as usual, and let $\mu = \zeta^{-1}$. Then:
\begin{align*}
f = g \zeta & \iff g = f \mu.
\end{align*}
Equivalently,
\begin{align*}
\forall x \in X, f(x) = \sum_{y \in X} g(y)\zeta(y,x) &\iff \forall x \in X, g(x) = \sum_{y \in X} f(y)\mu(y,x).
\end{align*}
\end{thm}
\begin{proof}
Recall that by Lemma \ref{inv_ex}, $\mu = \zeta^{-1}$ exists. Given that, the result follows if we right-multiply both sides by $\mu$. The first statement becomes the second when the sums are made explicit.
\end{proof}

This result is quite straightforward based on how we defined $\zeta$ and $\mu$. In fact, it feels strange that a supposedly central fact is just the inversion of a single matrix operation. In fact, the simplicity of the inversion step is the whole point of M\"obius inversion. The tricky part is formulating a given problem such that Theorem \ref{mob_inv} can be applied to it. We will see many cases where values of $f$ from the theorem are readily attainable, but values of $g$ are what we really want. The M\"obius function allows us to obtain those values. This procedure - the use of the M\"obius function to reverse the process of summing terms as in theorem Theorem \ref{mob_inv} - is what is referred to as M\"obius inversion.

@@

In this case, if $\leq$ represents the poset relation from Definition \ref{lattice}, then $m \leq n$ if, for each prime $p$, $m$'s factorization contains no more copies of $p$ than $n$'s factorization. This is the same as saying that $m|n$. So the poset relation becomes divisibility. We will explore this example more in the next sections.
\end{examp}
It turns out that several important posets have this lattice structure. So the goal is to determine the M\"obius function on certain subsets of the integer lattice, and then transfer those results to other partially ordered sets. To do so, we first need a lemma which applies to the M\"obius function on any poset.
\begin{lemma}\label{recurse}
Let $X$ be a lower finite partially ordered set, and let $\zeta, \mu$ be defined as usual. Then for all $a, b \in X$,
\[
\mu(a,b) = 
\begin{cases}
0 & a \nleq b\\
\delta(a,b)-\sum_{c \in [a,b)}\mu(a,c) & else.
\end{cases}
\]
\end{lemma}
\begin{rmk}
Note that the piecewise form is actually not necessary, because if $a \nleq b$ then the sum will be empty. So it is correct - albeit less clear- to just write
\[
\mu(a,b) = \delta(a,b)-\sum_{c \in [a,b)}\mu(a,c).
\]
\end{rmk}
\begin{proof}
We fix $a,b$, and then rearrange terms starting from Remark \ref{sum_form}:
\begin{align*}
\sum_{c \in X}\mu(a,c) \zeta(c,b) &= \delta(a,b)\\
\sum_{c \in [a,b)} \mu(a,c) + \mu(a,b) &= \delta(a,b)\\
\mu(a,b) &= \delta(a,b) - \sum_{c \in [a,b)} \mu(a,c) 
\end{align*}
\end{proof}
We are now ready to write down $\mu$ on the integer lattice.
\begin{thm}\label{int_lattice}
Let $L$ be the $n$-dimensional integer lattice, with the usual poset relation. Let $\zeta, \mu$ be defined as usual on $L$. Let $x=(x_1,\dots,x_n),y = (y_1,\dots,y_n)$ be elements of $L$. Then:
\[
\mu(x,y) =
\begin{cases}
0 & x\nleq y\\
0 & \exists i \in \Nn: y_i-x_i \geq 2\\
(-1)^{\sum_{i=1}^ny_i-x_i} & \text{otherwise}
\end{cases}
\]
\end{thm}

\begin{proof}
If $x \nleq y$, then by Lemma \ref{recurse}, $\mu(x,y) = 0$. This is the first case.

We now proceed by induction on $\sum_{i=1}^ny_i-x_i$. If $x \leq y$ and $\sum_{i=1}^ny_i-x_i = 0$, then $x = y$. So by Lemma \ref{recurse}, $\mu(x,y) = 1$. This is the base case.

Suppose that we have Lemma \ref{int_lattice} for any $x \leq y$ where $\sum_{i=1}^ny_i-x_i \leq k-1$. Let $x \leq y$ be fixed such that $\sum_{i=1}^ny_i-x_i =k$. Suppose that for some $j \in \Nn, y_j-x_j \geq 2$. Let $y' = (y_1,\dots,y_j-1,\dots,y_n)$. Note that $x \leq y' < y$. So we can decompose the sum from Lemma $\ref{recurse}$. Let
\begin{align*}
\alpha &= \sum_{x \leq z \leq y'} \mu(x,z)\\
\beta &= \sum_{y' < z < y}\mu(x,z).\\
\end{align*}
Then
\begin{align*}
\mu(x,y) &= -\sum_{x \leq z \leq y}\mu(x,z)\\
&= -\sum_{x \leq z \leq y'} \mu(x,z) - \sum_{y' < z < y}\mu(x,z)\\
&= -(\alpha + \beta).
\end{align*}
We note that
\[\alpha = \sum_{z \in L} \mu(x,z)\zeta(z,y').
\]
So by Remark \ref{sum_form}, $\alpha=0$.

@@

Finally, suppose that for all $i = 1,\dots,n, y_i-x_i \leq 1$. Because $\sum_{i=1}^ny_i-x_i =k$, there must be $k$ indices where $x$ and $y$ differ by one, while at the other $n-k$ indices they are equal. So for each $j = 0,\dots,k-1$, there are $k \choose j$ elements $z$ of $X$ where $x \leq z < y$ and $\sum_{i = 1}^n y_i-z_i = j$. For each of these, by the inductive hypothesis, $\mu(x,z) = (-1)^j$. So, using Remark \ref{sum_form}, we can write
\begin{align*}
\mu(x,y)& = -\sum_{j = 1}^k {k \choose j} (-1)^j\\
&= (-1)^k.
\end{align*}
\end{proof}
This result will be powerful for determining values of $\mu$ on various posets whose structure is similar to an integer lattice. 
\begin{thm}\label{mapping}
Let $(X, \leq_X),(Y, \leq_Y)$ be lower finite posets. Let $x_1 \leq x_2 \in S$ be given. Suppose that there exists a bijective map
\[
\phi: [x_1, x_2] \leftrightarrow [\phi(x_1), \phi(x_2)]
\]
such that for all $z_1, z_2 \in [x_1, x_2]$,
\[ 
z_1 \leq_X z_2 \iff \phi(z_1) \leq_Y \phi(z_2).
\] 
Let $\mu_X, \mu_Y$ denote the M\"obius functions on $X$ and $Y$ respectively. Then:
\[
\mu_X(x_1, x_2) = \mu_Y(\phi(x_1), \phi(x_2)).
\]
\end{thm}
This theorem tells us that if we can find a portion of a familiar poset which has identical structure to a portion of an unknown poset, we can transfer what we know about values of $\mu$ on that portion from one to the other.
\begin{proof}
We induct on $\#[x_1, x_2]$. Because $x_1 \leq x_2$, if $[x_1, x_2] = 1$, then $x_1 = x_2$. So by Lemma \ref{recurse}, $\mu_X(x_1, x_2) = 1$. Because $\phi(x_1) = \phi(x_2), \mu_Y(\phi(x_1), \phi(x)2) = 1$ also.

Suppose the theorem is true if $\#[x_1, x_2] \leq k-1$ for some integer $k$. Fix $x_1, x_2$ such that $\#[x_1, x_2] = k$. Then by Lemma \ref{recurse},
\[
\mu_Y(\phi(x_1), \phi(x_2)) = -\sum_{y \in [\phi(x_1), \phi(x_2)]}\mu_Y(\phi(x_1), y)
\]
Because $\phi$ is an isomorphism which preserves $\leq$, we can rewrite this sum:
\[
\mu_Y(\phi(x_1), \phi(x_2)) = -\sum_{x \in [x_1, x_2]}\mu_Y(\phi(x_1), \phi(x))
\]
We can restrict $\phi$ to $[\phi(x_1), \phi(x)]$ to satisfy the hypotheses for all intermediate elements $x$. So we can apply the inductive hypothesis to all the terms of this sum:
\begin{align*}
\mu_Y(\phi(x_1), \phi(x_2)) &= -\sum_{x \in [x_1, x_2]}\mu_X(x_1, x)\\
&= \mu_X(x_1, x_2)
\end{align*}
\end{proof}
The following corollary follows immediately if we let $\phi$ be the inclusion map:
\begin{cor}\label{subset}
Let $X$ be a poset. Let $I$ be an interval in $X$. In other words, let $I = [z_1, z_2]$ for some $z_1, z_2 \in X$. Then
\[
\mu_I = \mu_X|_I.
\]
\end{cor}
In the next section, we explore a poset where this fact will be extremely useful.

@@

We will consider the poset $(\Zz^{+}, |)$. This is the set of positive integers ordered by divisibility (recall that $d|n$ if $d$ divides $n$ without remainder). Note that this poset is lower finite: For any number $n$, the set $\{q \in \Zz^{+}: q|n\} \subseteq \{q \in \Zz^{+}: q \leq n\}$, and the latter set contains $n$ elements, which is finite (note that $d|n$ means that $d$ divides $n$ without remainder). This means we are able to use M\"obius inversion. Rather than compute the M\"obius function directly, we can use Theorems \ref{int_lattice} and \ref{mapping} to reuse what we already know.

@@

\begin{proof}
Let $d,n \in \Zz^+$ be given. We set up the isomorphism required for Theorem \ref{mapping}: Index the primes $p_1, p_2, \dots$. So $p_1=2, p_2=3$, and so on. Let $k$ be the index of the largest prime which divides either $d$ or $n$. Then let $\phi : [d,n] \rightarrow \Nn^k$ be defined by
\[
 p_1^{x_1}p_2^{x_2}\cdots p_k^{x_k} \mapsto (x_1, \cdots, x_k).
\]
We claim that $\phi$ is a bijection from $[d,n]$ to $\im(\phi)$.

@@

This allows us to use Theorem \ref{mapping}, and the result immediately follows.
\end{proof}

This definition of $\mu$ depends only on $\frac nd$, which means that we can abuse notation to write $\mu(\frac nd)$ instead of $\mu(d,n)$. This allows us to use Theorem \ref{mob_inv} to arrive at \eqref{init_mob}:
\[
f(n) = \sum_{d | n} g(d) \iff g(n) = \sum_{d | n}  f(d)\mu\left(\frac nd\right).
\]
\section{Euler's Totient} %TODO: Fix \Zz^+
Euler's Totient is a famous function in number theory, which counts the how many numbers are relatively prime to a particular value. We denote the function by $\varphi : \Nn \rightarrow \Nn$, and its definition is:
\[
\varphi(n) = \#\{1 \leq m \leq n : (m,n) = 1\},
\]
where $(m,n)$ denotes the greatest common divisor of $m,n$.

@@

We start with an important fact:
\begin{thm}[Gauss]\label{gauss}
Let $n \in \Nn$ be given. Then 
\[
n = \sum_{d|n}\varphi(d).
\]
\end{thm}
There are a variety of ways of proving this fact. Here we include a concise proof using group theory.
\begin{proof}
For any integer $d$, $\varphi(d)$ counts the generators of $C_d$. Each element of $C_n$ generates exactly one cyclic subgroup, so counting the total number of generators of cyclic subgroups of $C_n$ is equivalent to counting the elements of $C_n$. Also, $C_n$ has exactly one cyclic subgroup for each divisor of $n$ (and no others). So the sum indeed counts the number of elements of $C_n$, of which there are $n$.
\end{proof}
This sum gives us a value we have access to ($n$) in terms of a sum of values we want (values of $\varphi$). This is a perfect situation to apply M\"obius inversion, and undo that sum. We apply \eqref{init_mob}, which tells us
\[
\varphi(n) = \sum_{d|n} d\mu\left(\frac nd\right).
\]
Recall that Theorem \ref{mu_div} tells us how to write down values of $\mu$. Depending on what information we have access to, this might be much easier than just counting the number of values relatively prime to $n$. This is a straightforward example of how M\"obius inversion's ability to reverse a sum can be useful.
\end{chapter}

\begin{chapter}{Groups and Orbits} \label{chap_groups}
\section{Background}
For this chapter, it is necessary to be familiar with a simple structure called a \emph{group}. In this section, we will define it and briefly establish the machinery we need without going into too much detail, omitting proofs. More detail on elementary group theory can be found in an abstract algebra textbook. Knowledge of groups is not needed for any other chapters in this investigation.

\begin{defn}
A \emph{group} is a set $G$ together with a group operation (denoted with the empty operator) $G \times G \rightarrow G$. This operation must satisfy three requirements.
\begin{itemize}
\item For all $a,b,c \in G, (ab)c = a(bc)$
\item There exists an element $e \in G$, such that $ea=a=ae$ for all $a \in G$
\item For each $a \in G$, there exists an $a^{-1} \in G$ such that $aa^{-1} = e = a^{-1}a$.
\end{itemize} 
\end{defn}
\begin{defn}
Let $G$ be a group. Then $H \subseteq G$ is a \emph{subgroup} of $G$ if $H$ is itself a group, with the same group operation as $G$. If $H$ is a subgroup of $G$, we write $H \leq G$.
\end{defn}
\begin{defn}
Let $G$ be a group and let $\Omega$ be a set. Then an \emph{action} of $G$ on $\Omega$ is a map $\cdot:G \times \Omega \rightarrow \Omega$ which satisfies the following properties:
\begin{itemize}
\item $e \cdot x = x$ for all $x \in \Omega$
\item $g \cdot (h \cdot x) = (gh) \cdot x$ for all $g,h \in G, x \in \Omega$.
\end{itemize} 
\end{defn}
\begin{defn}
Let a group $G$ act on a set $\Omega$. Let $x \in \Omega$ be given. Then the \emph{orbit of $x$ under $G$}, denoted by $\orb_G(x)$ or $Gx$, is defined by
\[
\orb_G(x) = \{ y \in G : \exists g \in G \text{ where } g\cdot x = y\}
\]
or, equivalently,
\[
\orb_G(x) = \{g\cdot x : g \in G\}.
\]
\end{defn}
